High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-07-20 inst/manuscript/manuscript.Rmd
Fully revised manuscript (except abstract), SI, and figures.

Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019]. Significant efforts are being made to address the various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g. @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts. Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time, ideally on an annual basis, given the rapid pace of change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010]. Unfortunately, the data that do exist are in many places inaccurate and infrequently updated. Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013]. Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018]. Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016; @levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].

These informational inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]. Cropland mapping over Africa is difficult for several reasons. The primary reason relates to the characteristics of the continent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019]. This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g. @ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979; @boschettiAnalysisConflictOmission2004]]. In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons. These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.

Recent technological advances are helping to overcome these challenges. Chief among these are the growing numbers of satellites that collect high (\<5 m) to near-high (10 m) resolution imagery at sub-weekly intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. This high spatial *and* temporal resolution imagery addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, advances in cloud computing and the opening of satellite image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale [@GorelickGoogleEarthEngine2017]. These capabilities are already being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017; [@zhangGLCFCS30Global2021a]]. However, the potential of the highest resolution (\<5 m) imagery to map cropland over very large extents (e.g. country scales) has yet to be realized, presumably because these data are commercial and relatively expensive, and require significant computational resource to process.

Beyond the imagery and computational gains, machine learning algorithms are rapidly advancing, providing large gains in classification performance [@maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018]. However, the ability to take advantage of these gains is often limited by newer models' need for large training datasets, which are typically unavailable, hard to collect, or contain numerous errors in developing regions [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @burkeUsingSatelliteImagery2021]. To build sufficient samples for training cropland mapping models, as well as the reference data needed to objectively assess their performance (we refer collectively to both sample types as "labels", distinguishing between each type where needed), map-makers rely heavily on visual interpretation of high resolution satellite or aerial imagery [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019], as it is impractical and expensive to collect these data in the field over large areas, particularly on an ongoing basis. Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Visually interpreted labels present two particular problems. The first is that they inevitably contain errors of interpretation that can vary substantially according to the skill of the labeller, particularly over complex croplands with small field sizes [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. The second problem is that visual interpretation depends on high resolution imagery (\<5 m), as field boundaries become increasingly difficult to discern as image resolution decreases. Typically the only practical source of high resolution imagery is "virtual globe" basemaps (e.g. Bing or Google Maps), which are composed of mosaics of various high resolution satellite and aerial images that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which is usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017). If a land change occurs in the interval between the two image sets (e.g. a new field was created), the label, even if accurately drawn, introduces error into the classifier. This source of error may be elevated in tropical smallholder systems, where swidden agricultural is common [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
Despite the high potential for label error, they are typically not accounted for during model training and map accuracy assessment, resulting not only in the potential for maps to be misused or misinterpreted in subsequent analyses,  but in missed opportunities to improve model performance [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018; @StehmanKeyissuesrigorous2019].

Taking into consideration the advances and remaining limitations described above, the ability to map smallholder-dominated croplands can be further improved by 1) more fully exploiting the profusion of high frequency, high resolution imagery provided by CubeSats [@McCabefutureEarthobservation2017], and 2) by implementing methods that improve the ability to collect and minimize errors in image-interpreted labels. We developed a mapping approach that focuses on these two sources of improvement. Our approach uses PlanetScope imagery collected by Planet's fleet of Dove satellite, which provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017; @planetteamPlanetApplicationProgram2018], at relatively low to no cost for academic research\footnote{www.planet.com/markets/education-and-research/} and non-commercial, sustainability-oriented applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}. Although these data are of lower spectral depth and, in some cases, quality, than Landsat, Sentinel, or Worldview imagery, their daily revisit enables country- to continent-scale image mosaics to be created for multiple periods during a single agricultural year, even over the cloudiest forest regions where it is hard to successfully construct cloud-free composites from optical imagery with return intervals (even by a few days). This ability to capture intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016]. Beyond the frequency, PlanetScope's 3.7 m resolution--although substantially coarser the 0.5-1 m available in most areas covered by virtual globes--is sufficiently resolved for humans to discern small fields under many conditions [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be made using the same imagery that is classified, which helps to minimize label error. To further reduce label noise, we developed a platform that includes rigorous label accuracy assessment protocols and a novel approach for creating consensus labels, which helps reduce mistakes made by individual labellers [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. We couple the labelling platform with a machine learning model inside an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] framework, in which model training is done iteratively and interactively, wherein the model's prediction uncertainty over unlabelled areas is assessed after each training step, and a sample of the most uncertain sites is selected for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011]. This approach helps boost the performance of the classifier while reducing the overall number of labels required to achieve a given level of performance [e.g. @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to convert pixel-wise cropland predictions into vectorized maps of individual field boundaries.

Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, ranging from primarily grain and vegetable crop production in the northern savannas to tree crop-dominated systems in the forested southwest, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. The map represents a single agricultural year (2018-2019), as opposed to a multi-year epoch, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size. In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, we provide a new technique for converting daily imagery into seasonal composites, and show how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].

# Materials and Methods
The mapping approach we developed is comprised of four open source components (Figure \ref{fig:systemoverview}) that are designed to run in a cloud computing environment. The first component collects daily PlanetScope imagery and converts them into cloud-free seasonal composites. The second is a custom-built platform that provides tools for labelling the composites, along with procedures to assess and minimize label error. This platform interacts with the third component, a machine learning process, in an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] loop, to produce a map of predicted cropland probabilities for each image pixel. The fourth and final component is an algorithm that segments the image composites, then filters the resulting polygons using the pixel-wise cropland predictions produced by the active learning classifier, resulting in a final set of vectorized field boundaries.

@@

## Image compositing
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("image_quality")
data("tile_key")
tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
  group_by(name, season, worker_id) %>% 
  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() 
tilen <- tile_quality %>% distinct(tile) %>% count() %>% pull(n)
```
The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency. The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016] that are not fully captured by the provided cloud masks. To minimize the effect of these residual errors, we developed a procedure for creating temporal composites of the primary growing and non-growing seasons within a single 12-month period. For Ghana, we defined the primary growing season as May through September, followed by the off (or dry) season from November or December through February.
We chose these two seasons because prior work shows that the contrast improves cropland classifications [@Debatsgeneralizedcomputervision2016], while having the seasons captured in sequence in the same year helps minimize differences caused by land change. The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable (and thus unavailable in Planet's catalog), thus the frequency of available imagery can be substantially less than daily during several months.

We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (beginning in May and ending in February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3). We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each to the boundaries of a 0.05$^\circ$ grid that it intersected (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for making composite image tiles. We chose this cell size for tiling because it is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.

To create a seasonal composite tile, we calculated two weights for the time series of each pixel within the ARD stack for a given season:

@@

```{=tex}
\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}
```
Where *t* is a particular date in the pixel time series, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]

@@

We generated a catalog of `r nrow(tile_key) * 2` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana. To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category scoring system, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g. because of between-date image mis-registrations). Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), making the highest possible score 12. We rescaled scores to fall between 0 and 1.

## Mapping cropland probabilities with active learning 
The first step in creating a country-wide field boundary map of Ghana was to create a pixel-wise classification of cropland probabilities throughout the country. Given the high resolution of the imagery, we divided Ghana into 16 distinct mapping regions, or Areas of Interest (AOIs), to minimize the computational burden. We constructed the AOIs by grouping together tile grids into blocks representing the larger 1$^\circ$ cells used to assign tile grid identifiers (Figure S1A). We grouped tile grids from 1$^\circ$ degree cells that overlapped Ghana's boundaries together with tile grids from the adjacent 1$^\circ$ cell contained entirely within Ghana, with the exception was AOI 16, comprised of tile grids from the 1$^\circ$ cells intersecting Ghana's southern coast. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$).

We used the active learning process to develop a separate cropland classification model for each of these AOIs, based on an approach described by Debats et al (2017). We initiated the process by training a starter model using labels from a set of randomly selected training sites drawn from a 0.005$^\circ$ grid that was nested within the tiling grid. This finer grid, which we refer to as the "primary grid" for simplicity, provided the target area for creating labels (section 2.2.1), as well as the unit for distributing computing jobs (section 2.2.2). We then assessed the performance of the starter model against a separate set of validation labels developed for each AOI. The trained model was then applied to predict cropland probabilities for pixels in unlabelled primary grid cells in each AOI, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] was calculated:

@@

Where Q is the uncertainty for each unlabelled primary grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it. Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus the lowest values of $Q$ represent locations posing the most difficulty for the classifier.

We ranked the unlabelled primary grid cells in from least to most certain, and randomly selected a subset cells from the top 30% of the ranking (to minimize the risk of spatial autocorrelation) and sent these back to the labelling platform. After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was sent for labelling. This loop repeated until model performance gains saturated or reached a threshold, and then a final map of predicted cropland probabilities was made for each AOI.

>>>>

Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019]. A number of efforts are underway to address various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g. @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts. The success of these efforts depends heavily on data that accurately describes the location and characteristics of croplands [@FritzMappingglobalcropland2015], and, given the rapid pace of agricultural change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010; @ZengHighlandcroplandexpansion2018], how these are changing from one year to the next. Unfortunately, for many regions, existing cropland datasets are inaccurate, and are usually created as once-off or infrequently update products. As such, estimates of global cropland area tend to vary widely, often disagree about where croplands are located [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], and become rapidly outdated. Errors in these maps can propagate in subsequent analyses that use cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018]. Beyond distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016; @levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].

The deficit of information is due to the fact that in many regions the only source of cropland data are remotely sensed land cover maps, which are prone to error. This is particularly true in Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021], and where the characteristics of croplands exacerbate the error inherent in remote sensing analyses. Half of all fields in Africa's smallholder-dominated agricultural systems are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019]. This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g. @ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979; @boschettiAnalysisConflictOmission2004], wherein the pixel's shape does not match that of crop fields, and is too coarse to aggregate into an approximation of that shape [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is the high variability within and between fields, their tendency to intergrade with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and the high temporal variability within croplands. These last three aspects pose challenges for the classification algorithms that are applied to the imagery.

Recent technological advances are helping to overcome these challenges. Chief among these are the growing numbers of satellites that collect high (\<5 m) to near-high (10 m) resolution imagery at sub-weekly intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. The spatial resolution of these imagery addresses the scale mismatch between sensor and field, and their high frequency captures the seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, the opening of satellite image archives [@WulderglobalLandsatarchive2016] and advances in cloud computing are placing large volumes of moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale [@GorelickGoogleEarthEngine2017]. These capabilities have already been used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017; [@zhangGLCFCS30Global2021a]]. However, the potential of the highest resolution (\<5 m) imagery to map cropland over very large extents (e.g. country scales) has yet to be realized, presumably because these data are commercial and relatively expensive, and require significant computational resource to process.

Beyond the imagery and computational gains, machine learning algorithms are rapidly advancing, providing large gains in classification performance [@maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018]. However, the ability to take advantage of these gains is often limited by newer models' need for large training datasets, which are typically unavailable, hard to collect, or contain numerous errors [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @burkeUsingSatelliteImagery2021]. To build sufficient training samples, as well as the reference data needed to objectively assess their performance (we refer collectively to both types as "labels", distinguishing between each as needed), map-makers rely heavily on visual interpretation of high resolution satellite or aerial imagery [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019], as it is impractical and expensive to collect these data in the field over large areas, particularly on an ongoing basis. As such, a number of web-based platforms have been developed to collect such labels [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Image-drawn labels present two particular problems. The first is that they inevitably contain errors of interpretation, which can vary substantially according to the skill of the labeller, particularly over complex croplands with small field sizes [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. The second problem is that visual interpretation depends on high resolution imagery (\<5 m), as fields are increasingly difficult to discern as image resolution decreases. Typically the only available source of high resolution imagery is "virtual globe" basemaps (e.g. Bing or Google Maps), which present mosaics of high resolution satellite and aerial images collected over a span of several years [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can create a temporal mismatch between the labels and the imagery being classified, which is usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017). If a land change occurs in the interval between the two image sets (e.g. a new field was created), the label, even if accurately drawn, introduces error into the classifier. This source of error may be elevated in croplands where swidden agriculture is practiced [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
Despite the high potential for it, label error is often not considered during model training and map accuracy assessment, resulting not only in the potential for maps to be misused or misinterpreted, but in missed opportunities to improve model performance [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018; @StehmanKeyissuesrigorous2019].

Taking into consideration the advances and remaining limitations described above, the ability to map smallholder-dominated croplands can be further improved by 1) more fully exploiting the profusion of high frequency, high resolution imagery provided by CubeSats [@McCabefutureEarthobservation2017], and 2) by implementing methods that improve the ability to collect and minimize errors in image-interpreted labels. We developed a mapping approach that focuses on these two sources of improvement. Our approach uses PlanetScope imagery collected by Planet's fleet of Dove satellite, which provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017; @planetteamPlanetApplicationProgram2018], at relatively low to no cost for academic research\footnote{www.planet.com/markets/education-and-research/} and non-commercial, sustainability-oriented applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}. Although these data are of lower spectral depth and, in some cases, quality, than Landsat, Sentinel, or Worldview imagery, their daily revisit enables country- to continent-scale image mosaics to be created for multiple periods during a single agricultural year, even over the cloudiest forest regions where it is hard to successfully construct cloud-free composites from optical imagery with return intervals (even by a few days). This ability to capture intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016]. Beyond the frequency, PlanetScope's 3.7 m resolution--although substantially coarser than the 0.5-1 m imagery available in most areas covered by virtual globes--is sufficiently resolved for humans to discern small fields under many conditions [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be made using the same imagery that is classified, which helps to minimize label error. To further reduce label noise, we developed a platform that includes rigorous label accuracy assessment protocols and a novel approach for creating consensus labels, which helps reduce mistakes made by individual labellers [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. We couple the labelling platform with a machine learning model inside an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] framework, in which the model is trained interactively, using the model's prediction uncertainty over unlabelled areas to select new sites for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011]. This approach helps boost the performance of the classifier while reducing the overall number of labels required to achieve a given level of performance [e.g. @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017]. An unsupervised segmentation step is then applied to convert pixel-wise cropland predictions into vectorized maps of individual field boundaries.

Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates across a broad mix of climate and agricultural systems, ranging from primarily grain and vegetable crop production in the northern savannas to tree crop-dominated systems in the forested southwest, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. The map represents a single agricultural year (2018-2019), as opposed to a multi-year epoch, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size. In addition to providing valuable new data and insight into Ghana's agriculture, our study demonstrates one of the most spatially extensive agricultural applications of CubeSats to date, provides a new technique for converting daily imagery into seasonal composites, and shows how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].

# Materials and Methods
The mapping approach we developed is comprised of four open source components (Figure \ref{fig:systemoverview}) that are designed to run in a cloud computing environment. The first component collects daily PlanetScope imagery and converts them into cloud-free seasonal composites. The second is a custom-built platform that provides tools for labelling the composites, along with procedures to assess and minimize label error. This platform interacts with the third component, a machine learning process, within an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] loop, to produce a map of predicted cropland probabilities for each image pixel. The fourth and final component is an algorithm that segments the image composites, then filters the resulting polygons using the pixel-wise cropland predictions produced by the active learning classifier, resulting in a final set of vectorized field boundaries.

@@

## Image compositing
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("image_quality")
data("tile_key")
tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
  group_by(name, season, worker_id) %>% 
  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() 
tilen <- tile_quality %>% distinct(tile) %>% count() %>% pull(n)
```
The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency. The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016; @royGlobalAnalysisTemporal2021] that are not fully captured by the provided cloud masks. To minimize the effect of these residual errors, we developed a procedure for creating temporal composites of the primary growing and non-growing seasons within a single 12-month period. For Ghana, we defined the primary growing season as May through September, followed by the off (or dry) season from November or December through February.
We chose these two seasons because prior work shows that the contrast between them improves cropland classifications [@Debatsgeneralizedcomputervision2016], Furthermore, capturing the seasons in this sequence during the same year helps minimize differences caused by land change. The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable and therefore unavailable in Planet's catalog, thus the effective return interval can be substantially longer than 24 hours during the cloudiest months [@royGlobalAnalysisTemporal2021].

We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (defined here as March, 2018-February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3). We then converted each scene into analysis ready data [@DwyerAnalysisReadyData2018] by cropping each to the boundaries of a 0.05$^\circ$ grid that it intersected (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for making composited image tiles. We chose this cell size for tiling because it is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, thereby helping to minimize edge artifacts in the composites.

To create a seasonal composite, we calculated two weights for the time series of each pixel within the ARD stack for a given season:

@@

```{=tex}
\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}
```
Where *t* is a particular date in the pixel time series, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to these atmospheric features [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]

@@

We generated a catalog of `r as.integer(nrow(tile_key) * 2)` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana. To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image with resolution degraded below the 3.7 m PlanetScope resolution (e.g. because of between-date image mis-registrations). Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), making the highest possible score 12. We rescaled scores to fall between 0 and 1.

## Mapping cropland probabilities with active learning 
The first step in creating a country-wide field boundary map of Ghana was to create a pixel-wise classification of cropland probabilities throughout the country. Given the high resolution of the imagery and the need to minimize the computational burden, we divided Ghana into 16 distinct mapping regions, or Areas of Interest (AOIs). We constructed the AOIs by grouping together tile grids into blocks representing the larger 1$^\circ$ cells used to assign tile identifiers (Figure S1A). We grouped tile cells from 1$^\circ$ degree cells that overlapped Ghana's boundaries together with those from the nearest 1$^\circ$ cell contained entirely within Ghana (with the exception of AOI 16, which was comprised of tile grids from the 1$^\circ$ cells along Ghana's southern coast. The average extent of the resulting AOIs was 15,457 km$^2$ (range 12,160-23,535 km$^2$).

We used the active learning process to develop a separate cropland classification model for each of these AOIs, based on an approach described by Debats et al (2017). We initiated the process by training a starter model using labels from a set of randomly selected training sites drawn from a 0.005$^\circ$ grid that was nested within the tiling grid. This finer grid, which we refer to as the "primary grid" for simplicity, provided the target area for creating labels (section 2.2.1), as well as the unit for distributing computing jobs (section 2.2.2). We then assessed the performance of the starter model against a separate set of validation labels developed for each AOI, applied the model to predict cropland probabilities for pixels in unlabelled primary grid cells in each AOI, and calculated an uncertainty criterion [@DebatsIntegratingactivelearning2017]:

@@

Where Q is the uncertainty for each unlabelled primary grid cell I, calculated from the predicted probability *p* of a randomly selected subset of pixels (x, y) drawn from it. Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus the lowest values of $Q$ represent primary grid cells posing the most difficulty for the classifier.

We ranked the unlabelled primary grid cells from least to most certain, randomly selected a subset of cells from the top 30% of the ranking (to minimize the risk of spatial autocorrelation), and sent these back to the labelling platform. After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was again sent for labelling. This loop repeated until model performance gains saturated or reached a pre-defined threshold, after which a final map of cropland probabilities was made for the AOI.
